An Approach to Off-line Handwriting Recognition through Segmentation Techniques and Deep Neural Networks

### 2017-12-17 Report/paper.tex

\begin{algorithm}[tb]
    \caption{Contour combination algorithm}
    \label{alg:overlap}
\begin{algorithmic}
    \STATE {\bfseries Input:} $contour1$, $contour2$
    \STATE $xGap$ = calculateXaxisDistance $contour1$ $contour2$
    \STATE $yOverlap$ = calculateYoverlap $contour1$ $contour2$
    %TODO ::\IF {$xGap < averageWordWidth * (1 / (averageCharacterCount * 2)) $ {\bfseries and} $yOverlap > min(word1.height, word2.height) * requiredYaxisOverlap$}
    \STATE combineContours $contour1$ $contour2$
\end{algorithmic}
The chosen paramters are:
\\
$averageCharacterCount = 8$ 
\\
$requiredYaxisOverlap = 1/2$
\end{algorithm}

@@

\subsection{Skeletonization}
\label{sec:skel}
Next we skeletonize the word. This normalizes the thickness of the strokes to a single pixel and
allows us to create a vertical projection profile. The skeletonization algorithm used is the scipy adaptation of \cite{zsthinning}. % TODO reference scipy

@@

In this section we provide techniques to verify the correctness of segmentation points.
\cite{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures.
In this paper we discuss machine learning based approaches precedented by a few rule based techniques.

\subsubsection{Rule based techniques}
Most of the oversegmentation correction is done by our neural network.
However some simple preprocessing is added to remove obvious unwanted segmentation points.

We first remove removes the segmentation points found at the start and the end of the word image.
These segmentation points are an artifact of the used segmentation approach, as can be observed in Figure Figure \ref{fig:char_segm}.
These points can be removed without information loss, as the alphabet does not contain letters that entirely convert to a single pixel line after skeletonization.

Subsequently we group consequent segmentation points.
The occurence of these consequent segmentation points is another artifact from our segmentation approach, as can be seen in Figure \ref{fig:char_segm}.

From these groups of points, we take the center point as the new potential segmentation points.
However if one or more of the segmentation points in the group stems from empty collumns, the new segmentation point is chosen as the center point of the points representing the empty collumns in the word image.

Finally we aggregate groups of consequent potential segmentation points, separated by collumns containing less pixels than a certain threshold value $maxSumThresh$.
We have chosen a conservative value of $3px$ for this threshold.
This remedies the segmentation of slated lines inside characters, which can be an artifact of the skeletonization algirithm used.

A technique was tested using the feedback of the chracter recognition network to remove incorrect segmentation points.
However this turned out not to be ineffective, because the neural network tends to provide random probabilities with a high degree of certainty when images of multiple characters together were fed to the nerual network.
Since it is impossible to validate these responses from the neural network, this approach was abandoned and is not used in the final version.

More rule based filter could be added for more optimal processing of the potential segmentation points.
However the focus of our segmentation correction is the neural network used to remove incorrect segmenatation points.

@@

\section{Future work}
Our recognition system cannot be used for real-world applications. Every individual technique has a certain error, which increases the difficulty of the next step, e.g. if word segmentation contains mistakes, character recognition becomes much harder.
This results in a system which cannot correctly recognize handwritten text.
Many improvements are still possible.

@@

The character segmentation process is inadequate as it is, and generally unable to segment words adequately so that the character recognition step can recognise the found segments.
More postprocessing can be used through rule based filters and a more robust neural network to correct the found segmentation points, but this will never reach an optimal solution through rule based filters alone because of the ambiguity of this problem.

Finally, other approaches can also be considered.
State-of-the-art systems currently use segmentation-free approaches, this involves handwriting recognition without explicitly splitting words into characters. These methods are mostly based on hidden Markov models and neural networks.
This concept is explained and applied by \cite{presham}.
\section*{Acknowledgements}
We thank Sarah Vluymans for her comments on this paper.
\bibliography{paper}
\bibliographystyle{icml2016}

>>>>

\begin{algorithm}[tb]
    \caption{Contour combination algorithm}
    \label{alg:overlap}
\begin{algorithmic}
    \STATE {\bfseries Input:} $contour1$, $contour2$
    \STATE $xGap =$ calculateXaxisDistance $contour1$ $contour2$
    \STATE $yOverlap =$ calculateYoverlap $contour1$ $contour2$
    %TODO:\IF {$xGap < averageWordWidth * (1 / (averageCharacterCount * 2)) $ {\bfseries and} $yOverlap > min(word1.height, word2.height) * requiredYaxisOverlap$}
        \STATE combineContours $contour1$ $contour2$
\end{algorithmic}
The chosen paramters are:
\\
$averageCharacterCount = 8$ 
\\
$requiredYaxisOverlap = 1/2$
\end{algorithm}

@@

\subsection{Skeletonization}
\label{sec:skel}
Next we skeletonize the word. This normalizes the thickness of the strokes to a single pixel and
allows us to create a vertical projection profile. We use the skeletonization algorithm of \cite{zsthinning}.

@@

In this section we provide techniques to verify the correctness of segmentation points.
\cite{evalsplitpoints} discusses three approaches to evaluate segmentation point correctness: rule based techniques, cost function optimization based methods and machine learning based procedures.
In this paper we discuss machine learning based approaches precedented by a few rule based techniques.
An example of applying these correction techniques to an image is shown in \ref{fig:char_segm}.

\subsubsection{Rule based techniques}

First we remove segmentation points found at the start and the end of the word image.
These points can be removed without information loss, as the alphabet does not contain letters that entirely convert to a single pixel line after skeletonization.
Next consequent segmentation points are grouped. The center of these points is considered as segmentation point.
Unless if one or more of the segmentation points in the group stems from empty columns, the new segmentation point is chosen as the center of the points representing the empty columns in the word image.
Finally we aggregate groups of consequent potential segmentation points, separated by columns containing less pixels than a threshold value 3.
This remedies the segmentation of slanted lines inside characters.

A technique was tested using the output of the character recognition network to remove incorrect segmentation points.
This turned out to be ineffective because the neural network tends to provide high probabilities given images containing multiple characters. These probabilities are not meaningful and cannot be used as feedback for segmentation correctness.

@@

\section{Future work}
Our recognition system cannot be used for real-world applications. Every individual technique has a certain error, which increases the difficulty of the next step; e.g. if word segmentation contains mistakes, character recognition becomes much harder.
This results in a system which cannot correctly recognize handwritten text.
Many improvements are still possible.

@@

Finally, other approaches can also be considered.
State-of-the-art systems currently use segmentation-free approaches, this involves handwriting recognition without explicitly splitting words into characters. These methods are mostly based on hidden Markov models and neural networks. Due to the ambiguous nature of the word segmentation problem, a segmentation-free approach results in much better results.
This concept is explained and applied by \cite{presham}.
\section*{Acknowledgements}
We thank Sarah Vluymans for her comments on this paper.
\bibliography{paper}
\bibliographystyle{icml2016}
